A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data

### 2016-08-23 bioc-workflow.Rmd
A few updates and tl section

In addition, to matching names, the content of the feature metadata
for identical feature annotations must match exactly across the data
to be combined. In particular for these data, we expect the same
proteins in each replicates to be annotated with the same UniProt
entry names and descriptions, but not with the same coverage of number
of peptides or peptide-spectrum matches (PSMs).

```{r fvarnames}
fvarLabels(hyperLOPIT2015ms3r1)
fvarLabels(hyperLOPIT2015ms3r2)
```

@@

```{r qcbx, fig.width = 12}
par(mfrow = c(1, 2))
o <- order(hyperLOPIT$Fraction.No)
plotDist(hyperLOPIT[, o], pcol = "#00000010")
lines(colMeans(exprs(hyperLOPIT[, o])), col = "red", type = "b")
boxplot(exprs(hyperLOPIT[, o]), las = 2)
```

@@

```{r plotDist}
hlo <- hyperLOPIT[, order(hyperLOPIT$Fraction.No)]
plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
         pcol = "purple", fractions = "Fraction.No")
title(main = "Marker occupancy profiles along the gradient")
matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
         lty = 1, col = "darkblue", type = "l")
legend("topleft", c("Mitochondrion", "Peroxisome"),
       lty = 1, col = c("purple", "blue"), bty = "n")
```

@@

```{r pdtab, echo=FALSE}
i <- hyperLOPIT$Fraction.No %in% 7:10
pData(hyperLOPIT)[i, -3]
```

@@

```{r setColors, include=FALSE}
setStockcol(paste0(getStockcol(), 70))
```
```{r plotSVM, eval=TRUE}
## set point size of each protein to be inversely proportional to the
ptsze <- exp(fData(svmres)$svm.scores) - 1

@@

It is common when applying a supervised classification algorithm to
set a specific score cutoff on which to define new assignments, below
which classifications are set to unknown/unassigned. This is important
as in a supervised learning setup proteins can only be predicted to be
localised to one of the sub-cellular niches that appear in the
labelled training data. We can not guaranette (and do not expect) that
the whole class diversity to be represented in the labelled training
data as (1) finding markers that represent the whole diversity of the
cell is challenging (especially obtaining dual- and multiply-localised
protein markers) and (2) many sub-cellular niches contain too few
proteins to train on (we recommend a minimum of 13 markers per
sub-cellular class for stratified 80/20 partitioning and 5-fold
cross-validation). <!--Conversely, with defining markers one must also
be careful to not label too many markers, as this can lead to
overfitting and mis-assignment.-->

@@

@@

getMarkers(hl, fcol = "phenoDisco.Input")

opt <- knntlOptimisation(hl, gocc,
                         fcol = "phenoDisco.Input",
                         by = .5, 
                         times = 1,
                         xval = 5, k = c(3,3),
                         BPPARAM = par
  )
```

>>>>

In addition, to matching names, the content of the feature metadata
for identical feature annotations must match exactly across the data
to be combined. In particular for these data, we expect the same
proteins in each replicates to be annotated with the same UniProt
entry names and descriptions, but not with the same coverage of number
of peptides or peptide-spectrum matches (PSMs).
vfn
```{r fvarnames}
fvarLabels(hyperLOPIT2015ms3r1)
fvarLabels(hyperLOPIT2015ms3r2)
```

@@

```{r qcbx, fig.width = 12}
par(mfrow = c(1, 2))
o <- order(hyperLOPIT$Iodixonal.Density)
plotDist(hyperLOPIT[, o], pcol = "#00000010")
lines(colMeans(exprs(hyperLOPIT[, o])), col = "red", type = "b")
boxplot(exprs(hyperLOPIT[, o]), las = 2)
```

@@

```{r plotDist}
hlo <- hyperLOPIT[, order(hyperLOPIT$Iodixonal.Density)]
plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
         pcol = "purple", fractions = "Fraction.No")
title(main = "Marker occupancy profiles along the gradient")
matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
         lty = 1, col = "darkblue", type = "l")
legend("topleft", c("Mitochondrion", "Peroxisome"),
       lty = 1, col = c("purple", "blue"), bty = "n")
```

@@

```{r pdtab, echo=FALSE, eval = FALSE}
i <- hyperLOPIT$Fraction.No %in% 7:10
pData(hyperLOPIT)[i, -3]
```

@@

```{r setColors, include=FALSE}
setStockcol(NULL)
setStockcol(paste0(getStockcol(), 70))
```
```{r plotSVM, eval=TRUE}
## set point size of each protein to be inversely proportional to the
ptsze <- exp(fData(svmres)$svm.scores) - 1

@@

It is common when applying a supervised classification algorithm to
set a specific score cutoff on which to define new assignments, below
which classifications are set to unknown/unassigned. This is important
as in a supervised learning setup proteins can only be predicted to be
localised to one of the sub-cellular niches that appear in the
labelled training data. We can not guaranette (and do not expect) that
the whole class diversity to be represented in the labelled training
data as (1) finding markers that represent the whole diversity of the
cell is challenging (especially obtaining dual- and multiply-localised
protein markers) and (2) many sub-cellular niches contain too few
proteins to train on (we recommend a minimum of 13 markers per
sub-cellular class for stratified 80/20 partitioning and 5-fold
cross-validation - this allows a minimum of ~10 examples for parameter optimisation on the training partition i.e. 2 per fold for 5-fold cross-validation, and then 3 for testing the best parameters on the validation set). <!--Conversely, with defining markers one must also
be careful to not label too many markers, as this can lead to
overfitting and mis-assignment.-->

@@

The first step, as with any machine learning algorithm, is to
optimise any free paramaters of the classifier. For the *k*-NN TL
classifier these are the class weights, one per subcellular niche,
that control the proportion of primary and auxiliary data to use for
learning. A weight can take any real value number between 0 and 1. A
weight of $\theta = 1$ indicates that all weight is given to the primary
data (and this implicitly implies that a weight of $1 - \theta$ is given to
the auxiliary data, similarly a weight of $\theta = 0$ implies that all
weight is given to the auxiliary data (so 0 to the primary source).
If we conduct a parameter seach and test weights $\theta = {0, 1/3, 2/3, 1}$
for each class, and if we have say 10 subcellular niches, this
will result in `4^10` different combinations of parameters to test. The
parameter optimisation is therefore time consuming and as such
we recommend you make use of a computing cluster.

<!--In the example below
we first examine a PCA plot of the data and overlay the remaining
subcellular markers. We can see that the two ribosomes, two nuclear
compartments and plasma membrane are quite well-defined and separated
form the rest of the plot, indicating that they are well-resolved in
the hyperLOPIT experiments. For this reason, and in the interest of
parameter optimisation and time we have removed these classes from the
$k$-NN TL optimisation as we expect their weight to be 1.

- add code here

@@

opt <- knntlOptimisation(hl, gocc,
                         fcol = ,
                         by = .5, 
                         times = 1,
                         xval = 5, k = c(3,3),
                         BPPARAM = par
  )
```
-->
